---
title: "Data Science Workflows Course" 
subtitle: "HU OSIRIS: tlsc-dsfb26v-20_workflows"
author: "Alyanne de Haan, PhD; Bas van Gestel; Marc Teunis, PhD"
date: "11-4-2021"
output: 
  ioslides_presentation:
     widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducible Research

```{r, echo = FALSE, out.width=1000}
knitr::include_graphics(
  here::here(
    "images",
    "research_comic_phd.gif"
    )
  
)
```


## Course contents

 - Open Science (lessons 1)
 - Data management (lesson 2)
 - Using Git/Github (lesson 3 - 5)
 - Advanced RMarkdown (lesson 6 - 8)
 - Databases (lesson 9)
 
## Exercises & Course Exam

 - Each lesson has multiple Exercises
 - Some exercises do not have answers available
 - Some exercises need to be done together with you class mates
 - Do the group exercises in your Projecticum group
 - Each lesson has one or more portfolio assignments ('dagopdrachten')
 - Complete all portfolio assignments in separate Rmd files
 - Build a bookdown project from all these files
 - Host the bookdown on github, via github pages

## Requirements for the exam
 
 - You need to hand in **all assignments** to pass the course. You cannot compensate an assignment with another. 
 - Your portfolio is hosted on a public Github pages website
 - Your portfolio is a `{bookdown}` project
 - You portfolio is appealing and written in proper English
 - Your portfolio contains a short reading guide, introduction and table of contents
 - At the end of the portfolio, include a short reflection paragraphg
 
## Reflection and where to go next


 
 All information you need to perform these steps is included in the course reader
 
## Course reader

The course<sup>**$**</sup> reader can be found here:

https://datascience.hu.nl/rsconnect/dsfb2-workflows-reader/
 
The source<sup>**\***</sup> for the reader can be found here:

https://github.com/DataScienceILC/tlsc-dsfb26v-20_workflows
 
_\$This is the first time that we run this course_ <br/>
_\*You are invited to work together with us to improve the reader_

## Lessons

We planned long sessions to:

 - Enable you to spend time on the exercises and portfolio assignments
 - Seek interaction with your class mates and instructors
 - Tackle any technical issues

## Presence

 - We require you to be present during the online session to maximize interaction.  - We will keep track of your presence during this course.
 - Feel free to bring up issues or topics that are related to the course or the project work (Projecticum).
 - We kindly ask you to switch on your video when interaction with other course participants and instructors

## Link with Data Science Projecticum 

The _Data Science for Biology 2_ course consists of two parts:

 1. "Workflows for Data Science Course" (tlsc-dsfb26v-20_workflows)
 2. "Research Project - Data Science Projecticum (tlsc-dsfb26v-20_projecticum)
 
 You need to apply what you learned in workflows to your research project

## Questions and Issues

If you have questions during the online-sessions, reach out to your classmates or instructors. Instructors are available when online, or via the chat 

Technical issues or issues with the reader content? Create a new Github issue here:

https://github.com/DataScienceILC/tlsc-dsfb26v-20_workflows

## Good luck!

```{r, echo=FALSE, out.width=1000}
knitr::include_graphics(
  here::here(
    "images",
    "automated.gif"
  )
)
```
