---
title: "A minimal {ggplot2} example"
subtitle: "Code demo with {flipbookr}"
author: "Marc A.T. Teunis"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
```

---

```{r, mtcars_data, include=FALSE}
datasets::mtcars %>% 
  as_tibble()

```

`r chunk_reveal("mtcars_data")`

---
## Using `{SLOWggplot}`

 - SLOWggplot is a dialect of normal ggplot2 code. 
 - Build an image more gradually
 - Features and aesthetics are added incrementally
 - Ideal for this flipbook demo and learning ggplot2! 
 - See: https://evangelinereynolds.netlify.app/post/slow-ggplot/ for more info

---

```{r, mtcars_plot, include=FALSE}
mtcars %>%
  ggplot() + #BREAK
  aes(x = disp) +
  aes(y = mpg) + #BREAK
  geom_point(position = "jitter") + #BREAK
  aes(colour = as_factor(cyl))
```

`r chunk_reveal("mtcars_plot", break_type = "user")`

---
## Break types
There are a number of break types that help you customize the flipbook

---

```{r, mtcars_non_seq, include=FALSE}
mtcars %>%
  ggplot() + 
  aes(x = disp) +
  aes(y = mpg) + 
  geom_point(
    size = 3,
    alpha = 0.5, 
    colour = "blue") +   
  aes(shape = as_factor(cyl)) +
  ggtitle("An example plot using the infamous mtcars")
```


`r chunk_reveal("mtcars_non_seq")`

---


<!-- adjust font size in this css code chunk, currently 80 -->

```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}

code.r.hljs.remark-code{
  position: relative;
  overflow-x: hidden;
}


code.r.hljs.remark-code:hover{
  overflow-x:visible;
  width: 500px;
  border-style: solid;
}
```



---

# Notes: 

- This minimal flipbook assumes you are online when you build and view.  It is going to find the most recent remark.js info online.

- xaringan/remark.js slideshows are not self contained, single file slide shows.  Make sure you ship all the associated files that are created to display your flipbooks/slideshows.  I like to push .html and associated files to github repository and use Github Pages to host for beyond local sharing. :-)

