--- 
title: "Course Workflows [HU OSIRIS: tlsc-dsfb26v-20_workflows]"
author: "<authors>"
date: "`r Sys.time()`"
site: bookdown::bookdown_site
output:
  bookdown::html_book:
    theme: darkly
bibliography: [references.bib]
biblio-style: apalike
link-citations: yes
always_allow_html: yes
github-repo: "uashogeschoolutrecht/tlsc-dsfb26v-20_workflows"
description: "<course details>"
params:
  geo_id: NL BE
  month: February March April May June July August September October November December
  year: 2020
---

```{r include=FALSE, eval=FALSE}
# source(here::here("code", "get_rmd_dependencies.R"))
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')

```

```{r, include=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
```

# Course introduction



Lesson 1; reproducible research

## version control

## scrum: hoe moet dat en hoe gaan we dat deze cursus ondersteunen. scrum moeten ze toepassen op projecticum.

Lesson 2 

## Uitleg en opzetten eigen github

o basic git workflow in commandline
o Start github
o Push en pull
o clone
o koppelen Rstudio

git: toepassen: voorbeeld-cases, koppelen aan Rstudio
o branching per fix en merging, braches pushen
o backtracking?

git: toepassen: groepsproject met bestanden projecticum
o teamwork: hoe samenwerken met git (wijs hoofd merge requests aan)
o github merge requests

Rmarkdown: Uitleg en zelf doorwerken voorbeelden (let op aansluiting met eventuele rmarkdown in daur1)
o Overzicht Formats (pdf, websites, books, dashboards)
o Basic syntax

o website online zetten via github 6. Rmarkdown: toepassen: o Meer syntax o rapporten (tekstbestanden) genereren (reproducible reports) 7. Rmarkdown: toepassen: o opzetten eigen Rmarkdown-homepage met informatie over ontwikkeling van student tijdens deze cursus. (projecticum, opdrachten SQL en GIT deze cursus)

SQL: Uitleg en zelf doorwerken voorbeelden
o mySQL?
o syntax basics
o toepassing basics
SQL: toepassen:
o SQL binnen Rstudio
o voorbeeld-cases uit life sciences
SQL: toepassen
o op data projecticum (of indien ongeschikt zelf in te brengen casus) en online presenteren op Rmarkdown-homepage



## Databases 

[database samples](https://wiki.postgresql.org/wiki/Sample_Databases)


voor refrences:
https://www.r-bloggers.com/2013/07/r-and-postgresql-using-rpostgresql-and-sqldf/
https://www.r-bloggers.com/2016/02/using-postgresql-in-r-a-quick-how-to/
https://www.datacareer.de/blog/connect-to-postgresql-with-r-a-step-by-step-example/
https://db.rstudio.com/dplyr/
https://dbplyr.tidyverse.org/

=


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6, fig.height = 4,
                      warning = FALSE,
                      error = FALSE, message = FALSE)


image_dir <- here::here(
  "images"
)

```

## Purpose of this website

## Course contents & materials

<p style="font-size:18px">The complete source code for the webinars and all dependent data, and files can be found on: 
```{r, echo=FALSE, results='asis', warning=FALSE, error=FALSE, message=FALSE}
urls <- readr::read_csv(
  here::here("urls")
)

urls[1,1] %>%
  unlist() %>%
  as.character() %>%
  paste()

```


## Getting the materials
To compile this website locally, you can clone the website repository from 
```{r, echo=FALSE}
urls <- readr::read_csv(
  here::here("urls")
)

urls[1,1]

```

To compile the book-site locally run `bookdown::render(".")` from inside the cloned repo. This repository can also be used to access the course materials directly in RStudio during the workshops.

## Getting R and RStudio
During the course we will be working on a Cloud Computing Environment which provides web access to R and RStudio via a Virtual Machine. This machine runs a server edition of RStudio and has the latest R version and all the required packages available. In order to access the server you will need credentials, which you will recieve before the course starts. This is a convenient way to use R in a course and during the course we will only be using this environment. This is to ensure reproducibility and prevents a lot of trouble shooting from your side and the teacher's side.

If you want to use R and RStudio locally on your laptop (the teacher will not support this during the course), this is where you can download the software from:

[RStudio](https://www.rstudio.com/products/rstudio/download/) 

[R](https://cran.r-project.org/)

## The `{bookdown}` package
This website was created using the `{bookdown}` [@R-bookdown] package written by Yihui Xie. The package can be downloaded from CRAN. For more information see the [documentation](https://bookdown.org/yihui/bookdown/).

The `{bookdown}` package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```
