# RMarkdown - Output formats {#rmarkdownoutput}
```{r setuples, include=FALSE}
les <- 6
```

## Learning outcome:
After this lesson you will be able to

 1. Have an overview of different output formates for RMarkdown and render those formats.
 1. Setup github pages
 1. Create a bookdown project
 1. Render a bookdown project to a local website (aka html pages)
 1. Add the rendered site to your github pages

## Easy start: job interviews

Before we start diving into RMarkdown again, read the following blog post. While you can see that we (the teachers) may have a slightly different opinion in some aspects (tidyverse!), it gives a nice checklist for you to scan on any omissions in your general R knowledge.[click](https://www.springboard.com/blog/27-essential-r-interview-questions-with-answers/)

## Output possibilities

With Rmarkdown you can generate many different things. This reader was made in Rmarkdown. The DAUR1 capstone pdf? RMarkdown. The latest report Alyanne had to send to an ethical committee? Again RMarkdown. 

These websites are examples too: [click](https://holtzy.github.io/Pimp-my-rmd/), [click](https://www.r-graph-gallery.com/index.html).

Have a look around on the following website to get a feeling for what is possible with Rmarkdown.

- [https://r-markdown-gallery.com/index.html](https://r-markdown-gallery.com/index.html)
- [https://bookdown.org/yihui/rmarkdown/basics-examples.html](https://bookdown.org/yihui/rmarkdown/basics-examples.html)

The options include:

  - documents
    - pdf
    - Word
    - html (also: blogs)
    - handouts
    - books
  - interactive documents
    - html widgets
    - dashboards
    - shiny applications
  - presentations
    - slides (Beamer, ioslides, Slidy etc)
    - Powerpoint presentation

### Setting output format

You can specifically choose which output format to render to with the RStudio "Knit" button in the toolbar, or with:

```
rmarkdown::render("your_file_name.Rmd", output_format = "word_document")
```

But usually, you use the YAML header to control which output format(s) knitting will produce and even customize the options for each format separately.
To customize each output format, you can alter the YAML header output argument:

```
---
title: "My document"
output:
  html_document:
---
```

You can change **html_document** to for instance **pdf_document** or **word_document**.

Or customize HTML and PDF output both, this is useful if you want to render multiple types of output.:

```
---
title: "My document"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    fig_caption: false
---
```

<div class="question">

##### Exercise `r les` {-}

This exercise is adapted from [this one from r4ds here](https://r4ds.had.co.nz/r-markdown.html#exercises-71)

Download Lorem.Rmd [here](https://github.com/DataScienceILC/tlsc-dsfb26v-20_workflows/blob/main/Rmd/lorem.Rmd).

From this .Rmd, knit one new document for each of the three built-in formats: HTML, PDF and Word. How does the output differ? How does the input differ? (You may need to install LaTeX in order to build the PDF output — RStudio will prompt you if this is necessary.)
</div>

### Output options

The YAML header can contain a lot more settings for your output. For instance, you may want to add a table of contents (TOC) to you html page, in this example a floating one:

```
---
title: "My document"
output:
  html_document:
    toc: true
    toc_float: true
---
```

Other settings can just be added below html_document:

#### Tabbed sections {-}

You can organize content using tabs by applying the .tabset class attribute to headers within a document. This will cause all sub-headers of the header with the .tabset attribute to appear within tabs rather than as standalone sections. For example:

```
## Quarterly Results {.tabset}

### By Product

(tab content)

### By Region

(tab content)
```

#### Themes {-}

You learned how to specify you own formatting with css last lesson. But RMarkdown has some [preset themes](https://www.datadreaming.org/post/r-markdown-theme-gallery/) you can use as well. And of course, you can download packages with more themes.



```
---
title: "Stuff"
output:
  html_document:
    theme: united
```
---

<div class="question">

##### Exercise `r les` {-}

  A) What is a floating TOC?
  B) Use Lorem.Rmd again and add a TOC
  C) Give the TOC a set depth (toc_depth) and see what happens if you change it. (for instance `toc_depth = 2` can be set just as you set `toc_float: true` earlier.)
  D) Also try `code_folding: hide` and `number_sections: FALSE`.
  E) Google how to put two authors on one .Rmd, with affiliations. Put your and your labpartner's names as authors.
  F) Apply a theme you like, but not "united".
  G) try tabbed sections

</div>


## PDF 

Rendering pdf files does not differ much from html. Themes work as well. There are few things you can set in the YAML header for pdf files. Note that these options do not get an indent, but rather need to be defined at the same level as title, author, output etc.

```
---
title: "My document"
output: pdf_document
fontsize: 11pt
geometry: margin=1in
---
```

## Word ...

Now word files are a different beast... You can export them easily enough:

```
---
title: "My weekly report"
author: Joep
date: March 22, 1999
output: word_document
---
```


But now how do we edit the layout? RMarkdown can use a reference .docx file to get all the layout information from.

<div class="question">
##### Exercise `r les` {-} 
 - Follow [these steps until "Beyond the basics".(click)](https://rmarkdown.rstudio.com/articles_docx.html)
 - Get your word-styles-reference-01.docx to work as a style reference.

</div>


<!-- What if some collaborator adds comments or tracked changes? -->
<!-- For now, nothing. I will try to get redoc to work -->

## Interactivity {#rmarkdowninteractive}

Until now, we discussed static output formats. But RMarkdown files can also render interactive content! There are two main types of content to render: html widgets and shiny apps. Shiny apps will be covered in two masterclasses in the projecticum. Shiny uses actual R code to render the interactivity. This gives you a very broad range of options, but also means that it requires a live R session running somewhere.

HTML widgets use JavaScript libraries to produce for instance interactive graphs, using R syntax! These will work without an R session, in any web browser.They will still only work in html though, you can't suddenly make pdf files interactive.

### Tables
We actually already showed you how to make interactive tables with `datatable()` in a previous lesson:

```{r, message=F, error=F, warning=F,}
library(DT)
datatable(iris, options = list(pageLength = 5))
```

### Javascript packages

You can do similar interactivity with graphs and figures. Generally, these packages link R to a javascript packages. Check the possibilities [here](http://www.htmlwidgets.org/showcase_leaflet.html). These packages actually do most of the work for you.

Let's try a few:

#### Maps {-}

The leaflet package let's you use maps (click on the dot and try moving the map)

```{r, message=F, error=F, warning=F, fig.cap = "You are here! But not in 2021, in 2021 we are at home"}
library(htmlwidgets)
library(leaflet)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lat=52.084036170323024, lng=5.173630727235891, popup="Here be HL7!") 
# get coordinates from google maps

m  # Print the map
```

<!-- You can change map colours if you like. type `names(providers)` for all the options. -->
<!-- No you can't in bookdown see https://github.com/rstudio/bookdown/issues/15 -->


Suppose you want to show all locations where you catched rabbits for your ongoing study on rabbit ear length at the Uithof:

```{r, message=F, error=F, warning=F, fig.cap="Here be rabbits"}
library(leaflet)
library(tidyverse)

rabbit_locations <- tibble( rabbitnr = c(1:5),
                            lat  = c(52.0852,52.0832,52.08247,52.0834,52.08309),
                            long = c(5.171109,5.171248,5.171248,5.176055,5.17268) )

leaflet(data = rabbit_locations) %>% addTiles() %>% 
  addMarkers(~long, ~lat, popup = ~as.character(rabbitnr), label = ~as.character(rabbitnr))
```

Want more? Check out this website [here](http://rstudio.github.io/leaflet/)

<div class="question">
##### Exercise `r les` {-} 

**rabbits!**

Change the markers in the example above to little icons of rabbits! See [here (click)](http://rstudio.github.io/leaflet/markers.html) how to do this.
</div>

#### 3d figures {-}

Sometimes 2 axes don't cut it. 

```{r, message=F, error=F, warning=F, fig.cap="With 3 axes you may want to be able to rotate the graph."}
library(threejs)
library(palmerpenguins)
z <- penguins$bill_length_mm
x <- penguins$flipper_length_mm
y <- penguins$body_mass_g
N = length(levels(penguins$species)) # we want 3 colours
scatterplot3js(x,y,z, color=rainbow(N)[penguins$species])


```

See some more examples, if you like, here: [click](https://bwlewis.github.io/rthreejs/)

<!-- <div class="question"> -->
<!-- ##### Exercise `r les` {-}  -->

<!-- **principal components of genomic variant data** -->
<!-- http://bwlewis.github.io/1000_genomes_examples/PCA.html#fn1 -->
<!-- Voor volgend jaar, memory issues -->
<!-- kleinere vcf here https://github.com/vcflib/vcflib/blob/master/samples/1kg-phaseIII-v5a.20130502.genotypes.chr22-16-16.5mb.vcf.gz -->

<!-- of andere les hier: https://broadinstitute.github.io/2020_scWorkshop/feature-selection-and-cluster-analysis.html -->

<!-- </div> -->

#### Time series {-}

This rabbit ear experiment has been going for quite a while. Let's say you started catching rabbits about 10 years ago and calculated average ear length every three weeks. The graph below shows the average ear length per 3 weeks. It also shows the time frame of the mysterious long ear disease (LED) outbreak of 2018:

```{r, message=F, error=F, warning=F, fig.cap="Average rabbit ear length at the uithof per 3 weeks for the last 10 years."}

# Libraries
library(dygraphs)
library(xts) 

# get some data
# Your time column MUST be a time format or numbers
rabbitdata <- data.frame(
  time=seq(from=Sys.Date()-(10*365), to=Sys.Date(), by=21 ), 
  oorlengte=runif((10*365)/21+1,min = 6, max = 10)
)

rabbitdata$oorlengte[120:138]<-rabbitdata$oorlengte[120:138]+4

# create xts format because dygraph() only eats xts
rabbitdata.xts <- xts(x = rabbitdata$oorlengte, order.by = rabbitdata$time)

# plot
dygraph(rabbitdata.xts) %>% 
  dySeries(label="average ear length (cm)", color="black") %>%
  dyShading(from=rabbitdata$time[120], to=rabbitdata$time[138], color="#FFE6E6") %>%
  dyRangeSelector(dateWindow = c(rabbitdata$time[90], rabbitdata$time[150]))


```

See some more examples, if you like, here: [click](http://rstudio.github.io/dygraphs/)

#### Plotly {-}
And most famously, [plotly](https://plotly.com/r/). This package can turn any ggplot2 graph into an interactive plotly graph. Mousing over a point displays information about that point. Clicking on a legend point removes that class from the plot. Clicking on it again returns it.

<!-- ####  rCharts -->
<!-- https://bookdown.org/BaktiSiregar/data-science-for-beginners/Advanced-Visualizations.html#interactive-graphs -->
<!-- shiny..hier nog maar niet, voor de masterclass -->

Plotly can just make your scatterplot zoom-able with labels for the data points:


```{r, message=F, error=F, warning=F, fig.cap="Minimal example plotly graph"}
library(plotly)

fig <- plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length)

fig
```

```{r}
#box plots
plot_ly(                  #creates first box plot
  y = rnorm(50),          #choose numeric variable for first plot
  type = "box") %>%       #specifies plot type is box plot
  
  add_trace(              #creates second box plot
    y = rnorm(50, 1))     #choose numeric variable for second plot
```


Basically, if you think "Hey, I have this ggplot graph and it is nice, but it could be clearer if it had some interactive options..." take a look at the [plotly examples here](https://plotly.com/r/). It is also rather useful when you want subplots in your figure, like this:

```{r, message=F, error=F, warning=F, fig.cap="Plotly with subplots. You can also use cowplot()"}
library(plotly)
fig1 <- plot_ly(economics, x = ~date, y = ~unemploy)
fig1 <- fig1 %>% add_lines(name = ~"unemploy")
fig2 <- plot_ly(economics, x = ~date, y = ~uempmed)
fig2 <- fig2 %>% add_lines(name = ~"uempmed")
fig <- subplot(fig1, fig2)

fig
```

It can do Sankey diagrams. Sankey diagrams are what the cool kids do nowadays, it seems. A Sankey Diagram is a visualisation technique that allows to display flows. 

```{r, message=F, error=F, warning=F, fig.cap="Somehow, plotly doesn't like our reader. The ugly lines shouldn't be there."}
fig <- plot_ly(
    type = "sankey",
    orientation = "h",

    node = list(
      label = c("A1", "A2", "B1", "B2", "C1", "C2"),
      color = c("blue", "blue", "purple", "purple", "darkred", "darkred"),
      pad = 15,
      thickness = 20,
      line = list(
        color = "black",
        width = 0.5
      )
    ),

    link = list(
      source = c(0,1,0,2,3,3),
      target = c(2,3,3,4,4,5),
      value =  c(8,4,2,8,4,2)
    )
  )
fig <- fig %>% layout(
    title = "Basic Sankey Diagram",
    font = list(
      size = 10
    )
)

fig
```


<!-- ### Flex Dashboards -->
<!-- I don't like them -->


## Bookdown {#bookdownportfolio}

You have generated a couple of .Rmd files for your portfolio by now. The next task will be to bundle them into one portfolio website, hosted by github.

The `{bookdown}` R package contains functions to bundle a collection of RMarkdown files in a website or a book, or a similar collection of information. It is based on RMarkdown and it's outpyut options (html, pdf, etc), but adds the possibility to make multi-page html websites and books sections, as well as cross-referencing across different .Rmd files. This website is rendered with bookdown. See some examples [here](https://bookdown.org/). Bookdown is used for any document that consists of multiple pages that are meant to be read in sequence, so books, readers, but also handouts, a thesis, graduation paper, a software manual, etc. And it *can* very well be used for multipage stuff that is not necessarily read sequentially, such as websites or a portfolio.


<div class="question">
##### Exercise `r les` {-}

**A simple bookdown example**

To start with bookdown, there is a [good and free resource available](https://bookdown.org/yihui/bookdown/get-started.html). 

 -Download the example zip of the [minimal example here: click](https://github.com/yihui/bookdown-minimal)
 -follow the steps in this link [(click)](https://bookdown.org/yihui/bookdown/get-started.html) and render the book.
 -If you don't have a `build book` button in your build tab, make sure you are in the demo project!
 
 ```{r, echo=FALSE, message=FALSE, fig.width=4 }
knitr::include_graphics(here::here(
    "images",
    "build_book.png"
  )
)
```
 

</div>

A bookdown book will consist of different chapters, each in their own .Rmd file (this is why you had to do this for your portfolio). The home page, or first page, is always index.Rmd, which contains the YAML header.

Chapters are rendered in alphabetical + numerical order of their .Rmd file name. So if you want 3 chapters, you will have files like `ìndex.Rmd`, `01_introduction.Rmd`, `02_cv.Rmd` and `03_firstproject.Rmd`. Go ahead and check the files in the repo of this course if you like, and compare them to the reader online.

File names that start with an underscore are skipped (in case you want to include files in the repo but not in the book for some reason).

Each R Markdown file must start immediately with the chapter title using the first-level heading, e.g., `# Chapter Title`. So chapters will look somewhat like this:

```
# About me

Stuff

## Paragaph title

More stuff

## Paragraph I don't want numbered {-}

bla

### subparagraph be here

more bla

```

<div class="question">
##### Exercise `r les` {-}
 - add two new pages to the example bookdown project.
 - render the bookdown again (build book)
</div>


### bookdown::render_book()

The `build book` button is calling a function: bookdown::render_book(). You need to check some boxes before this actually works. If you want to use the `build book` option in the future:

 1. Configure the build tools: Click on the menu button "Build" and then select "Configure build tools". You want to select "Website".
 1. Put this line in the YAML header: `site: bookdown::bookdown_site`

But you know our opinion on having to click anything at all, so you can use the console instead, for instance:

```
bookdown::render_book(here::here("index.Rmd"), "bookdown::gitbook")
bookdown::render_book("index.Rmd", "bookdown::pdf_book")
```

or simply try (really, try it. does it work?)

```
bookdown::render_book(".")
```

These options work if your chapter .Rmd files are in the project root directory (the one you see when you type `here::here()` in the console.)

If you really want to keep your chapter files in a subdirectory, which we don't recommend, you can still render the book with:

```
xfun::in_dir('mapje/nogeenmapje', bookdown::render_book('index.Rmd'))
```


<div class="question">
##### Exercise `r les` {-}

**Transform your portfolio to a local bookdown site**

By now you should have collected several Rmd files for completing the portfolio assignments in the course thusfar. If everything went well, you have these files collected in your portfolio RStudio project, linked to a Github repo.

 - Using the minimal example bookdown as a template, make a folder (locally) that contains a index.Rmd file that will act as the home page for your portfolio.
 - Start by rendering your project using `bookdown::render_book()` to see if it works.
 - Add the portfolio .Rmd files you generated in previous lessons to the folder. You can arrange the chapters by using file names like `001_<file_name>.Rmd`. 
 - When you have build your `index.Rmd` file and have collected all files in the root of your project, be sure all depending data, scripts and other files such as images etc. are also present in the project. Use the guerrilla analytics logic to setup the project structure if you lost track of it. 
 - Set the option `new_session: yes` in your `_bookdown.yml` or `index.Rmd` file. This prevent problems down the line. 
 - Render the book until all errors are solved. You can debug each file and error by looking at the line numbers where the rendering is causing problems.

</div>

## Hosting the bookdown project in your github pages

**IMPORTANT NOTE ON CONFIDENTIALITY**

<mark>**When hosting you portfolio on a public web address, please make sure that there is no confidential information visable in your project. Data that falls under a non-disclosure agreement for example from your projecticum may not be shared. Also, be sure not to publish any personal secrets such as passwords and tokens, inside a github repo or webpage.**</mark> 


### Github pages
Each account on github can be associated to one or multiple websites that are hosted on Github.com. To setup a website on Github (also called `github-pages` or `gh-pages`), we need to create a repository with a special name. For this course we will require you to upload results from the portfolio assignments to you personal github pages. **This is a hard requirement for passing this course. If you have no or an empty portfolio, you cannot pass this course** You may add any Rmd's with the general exersizes in this course if you feel like you would like to showcase them in your portfolio.



## Portfolio assignment `r paste0(les, ".1 ")` {-}  
<div class="dagopdracht">

**github pages**

 - Execute the steps explained [here (click)](https://guides.github.com/features/pages/) to setup a github-pages repository called `myname-portfolio` or something similar.
 
Once you have successfully rendered your bookdown to a full website, hosting can be achieved by copying all html files to your Github-pages. The easiest way is to:

 1. clone your github pages repo 
 1. copy all the html files rendered with bookdown to this folder
 1. commit and push all changes to your github pages repo. The website will be automatically updated after some time. 
 
You can also setup a theme for your github pages. See [here](https://docs.github.com/en/github/working-with-github-pages/adding-a-theme-to-your-github-pages-site-with-the-theme-chooser) for more info, but you have learned to use css, so this is a nice opportunity to display that.

 - Put your portfolio on your github pages
 - Send the url to your public portfolio address on the "prikbord" channel. 

</div>


## Portfolio assignment `r paste0(les, ".2")` {-}  
<div class="dagopdracht">

**looking ahead**

In class, we discussed this assignment yesterday. In the following weeks, start looking for what you will want to spend your extra time on this semester.

Try, for yourself, to answer the following questions:

 - Where do I want to be in ~2 years time?
 - How am I doing now with respect to this goal?
 - What would be the next skill to learn?

Make a planning on how to start learning this new skill. 

Note the following points:

 - It has to be life sciences or useful for life sciences (no knitting, riding horses etc)
 - It has to be data science / bioinformatics
 - You have to show your plan in your portfolio, together with the result of the first few steps you took in trying to rech this goal.
 - You probably won't learn this skill in a few days. Just get started.

This assignment will show us whether you succeeded in reaching an overview of the field and possibilities in data science for biology.

</div>

### More information
For a complete manual on working with github pages, [see the gh-pages docs](https://docs.github.com/en/github/working-with-github-pages/getting-started-with-github-pages) 

If you want to go for a full continuous integration workflow (which is highly recommended), you can follow these steps in [this blog](https://medium.com/@delucmat/how-to-publish-bookdown-projects-with-github-actions-on-github-pages-6e6aecc7331e)


