# Masterclass 2 - Shiny, building dashboards {#shinydashboards}

```{r, setuples, include=FALSE}
les <- 11
```

### Creating dashboards with shiny
Once you master the basics of shiny, you probably want to be able to build more complicated apps. One way of achieving a more visually appealing app and extend its functionality is by building a dashboard. A dashboard usually has some data input and multiple categories of information displayed from this data input. These categories are usually defined by tabs in the dashboard. To learn more about how to build nice dashboards in R go over [this tutorial:](https://rstudio.github.io/shinydashboard/get_started.html). You will need this information to complete the two portfolio assignments of this masterclass. 

<div class="dagopdracht">

##### Portfolio assignment `r les`

**Building a small dashboard for your projecticum**

For this assignment you need to build and publish a shiny dashboard. You need to use data from your projecticum assignment. We will organize a moment in class where you can show what you build. Because some of the projecticum projects are confidential, we ask you **not to publish** this app somewhere. The demo you will need to give will be done by running the app in a local (or on the Rstudio Server if you want) Rstudio session.   

For this assignment you will have to complete these steps:

 1. If you have not already done so, watch the [video tutorial on shiny](https://shiny.rstudio.com/tutorial/#video-tutorials)
 1. Go over the [`{shinydashboard}` tutorial](https://rstudio.github.io/shinydashboard/get_started.html)
 1. From the `{shinydashboard}` tutorial example build a small dashboard (with at least two tabs), using a dataset of your projecticum.
 1. Prepare for a short demo of your dashboard in the upcoming plenary class discussion. You will need to show your app and explain what it does and how your programmed it that way. 

</div>

### Interactive visualizations
Building a shiny app in which you or someone else can interact with the data ads a lot of interactivity to data. You can enable users to explore and visualize patters in the data and run multiple statistical models, without the need for a user to be able to program in R. Adding interactivity to a plot can also increase the user experience. One of the frameworks that is integrated with shiny is `{ggiraph}`. In our next assignment we will add some `{ggiraph}` elements to the dashboard you created for in the previous assignment.   
The `{ggiraph}` package has a number of build-in examples. Just like the shiny examples you can run them using the `run_girafe_example()` `{ggiraph}` function. To list all examples run

```{r, eval=FALSE}
library(ggiraph)
?run_girafe_example
```

Let's look at a `{ggiraph}` + `{shiny}` example

```{r, eval=FALSE}
run_girafe_example("crimes")
```

In this example you see how a 'normal' gg2plot is transformed into an interactive graph using the `renderGirafe()` function in the `server` part of the shiny app and the `girafeOutput()` function in the `ui` part of the app. Using these examples, you can easily incorporate an interactive `{ggiraph}` object in your own shiny dashboard. This is the aim of the next and final assingment of this lesson.  

<div class="dagopdracht">

##### Portfolio assignment `r les`

**Refactoring an existing shiny app to a dashboard and adding functionality**

Go over the following steps to complete this assignment:

 1. Go over the [`{ggiraph}` tutorial here:](https://davidgohel.github.io/ggiraph/articles/offcran/using_ggiraph.html) 
 1. Study the ["Shiny usage" tutorial part here:](https://davidgohel.github.io/ggiraph/articles/offcran/shiny.html)
 1. In the shiny dashboard you created in the previous \@ref(shinygettingstarted) masterclass assignment, add a `{ggiraph}` interactive element to this dashboard
 1. Add an additional dashboard tab where you plot covid-case data on the continent level - _maybe use a map to plot the data in tab 2 of the country in tab 1 on how that country is doing in relation to the rest of the coninent...?_  
 1. Update the dashboard code and push it to your projecticum github repository
 1. Share the link of the app-code with the teacher
 1. The teacher will deploy your app on a server to see it in action in a public url
 1. The best app (no correspondance to the jury ;-0), will recieve an endorsement in the ILC Today newsletter.
 
</div>

<div class="rstudio-tip">
 
To learn more about creating maps see these resources:

 - https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html
 - https://geocompr.robinlovelace.net/
 
 
</div>
 
 
 
