# Resources 


## Credits

### General resources

 - Wickham and Grolemund, "R for Data Science", 2017, O'Reily
 - Field et. al., 2015, "Discovering Statistics Using R", Sage Publications Ltd.
 - Sr. David Spiegelhalter, "The Art of Statistics", 2019, Pelican Books
  
## License resources

 - [Choose a License](https://choosealicense.com/). This website helps you to choose the best license for your project. 
 - [Using licenses on GitHub](https://docs.github.com/en/github/creating-cloning-and-archiving-repositories/licensing-a-repository). This GitHub documentation page provides information about how to add and apply licenses to your GitHub repositories.
  
## Git resources

### Git command line

 - [Git - the simple guide](http://rogerdudler.github.io/git-guide/). This web resource provides a simple explanation of Git and its commands.
 - [Pro Git book](https://git-scm.com/book/en/v2). This book explains the principles behind Git and also provides a great introduction to the Git command line. 

### Git and R(Studio)

 - [Collaboration and time travel using Git](https://rstudio.com/resources/webinars/collaboration-and-time-travel-version-control-with-git-github-and-rstudio/). In this video (approximately 45 minutes), R guru Hadley Wickham explains the concepts of Git collaboration and Git version  control. 
 - [Happy Git with R book](https://happygitwithr.com/). This book is a nice resource for R users that want to use Git in their projects.

## Github / SCRUM

 - https://github.com/features/project-management/
 - https://github.com/jvandemo/github-scrum-workflow
 - https://unito.io/blog/github-projects-agile/
 - https://dev.to/ben/for-those-who-make-use-of-github-projectswhats-your-process--5f7c [practices]
 - https://www.youtube.com/watch?v=idZyqNIrt84 [github automation and projects]
 - https://www.youtube.com/watch?v=YxKhb3fxtsU [github automation and projects]

## Youtube
 - https://www.youtube.com/watch?v=rz3_FDVt9eg 
 - https://www.youtube.com/watch?v=go5Au01Jrvs
 - https://www.youtube.com/watch?v=7UlWJWfZO9M
 - https://www.youtube.com/watch?v=GyNqlOjhPCQ
 - https://www.youtube.com/watch?v=1GmQ5BdAhG4
 
## R-bloggers

 - https://www.r-bloggers.com/how-to-learn-r-2/
 - https://digitheadslabnotebook.blogspot.com/2011/06/drawing-heatmaps-in-r.html?m=1
 

## Bioconductor
 
 - http://bioconductor.org/install/
 - http://bioconductor.org/help/package-vignettes/
 - http://bioconductor.org/packages/release/BiocViews.html#___Workflow

## Cheatsheets

 - https://www.rstudio.com/resources/cheatsheets/
 - https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf
 - https://www.rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf
 - https://github.com/rstudio/cheatsheets/raw/master/data-import.pdf

## Books 

 - http://r4ds.had.co.nz/
 - https://www.tidytextmining.com/
 
### Github

 - https://github.com/jennybc
 - https://github.com/hadley


## Github resource list

There are many 'R resources lists' to be found on the internet. Here, I just mention three that I liked.

 - https://paulvanderlaken.com/2017/08/10/r-resources-cheatsheets-tutorials-books/
 - https://www.r-project.org/doc/bib/R-books.html
 - https://github.com/oscarbaruffa/BigBookofR

## Python in RStudio

 - https://support.rstudio.com/hc/en-us/articles/360023654474-Installing-and-Configuring-Python-with-RStudio

## Databases

 - https://www.r-bloggers.com/2013/07/r-and-postgresql-using-rpostgresql-and-sqldf/
 - https://www.r-bloggers.com/2016/02/using-postgresql-in-r-a-quick-how-to/
 - https://www.datacareer.de/blog/connect-to-postgresql-with-r-a-step-by-step-example/
 - https://db.rstudio.com/dplyr/
 - https://dbplyr.tidyverse.org/
 - [database samples](https://wiki.postgresql.org/wiki/Sample_Databases)

## old Mendeley workflow for citations

For this example, we will use Mendeley. Mendeley is a free reference manager. If you are not yet using a reference manager, you start doing so right now. It will save you a lot of time in the future.

Go the [the Mendeley website](https://www.mendeley.com/download-desktop-new/), download and install Mendeley and open the program. (note: if you already use a reference manager, check if it provides BibTeX output. We need a .bib file later.)

If this is your first time using Mendeley: you can simply drag and drop pdf files into Mendeley. It will automatically start searching the web for the meta data. If your pdf is a published paper, there will be a lot of meta data available. 


#### How to get Mendeley into RMarkdown {-}
Special thanks to [Rosanna van Hespen](https://rosannavanhespen.nl/rmarkdown/writing-your-thesis-with-r-markdown-2-text-citations-and-equations/) for sharing her workflow online!

  1. First, start a new .Rmd file.

  2. Next, we need a file with all the references Mendeley is currently managing for you. At least 1 random paper and 1 self-made pdf. Mendeley can make a file containing this information: a .bib file. You can have Mendeley automatically updating this file for you, but as we are already syncing stuff with github, we will not make things this complicated for now. **So, you will need to repeat this step if you add new papers to Mendeley.**
  
  - Select the citations you want to include 
  - Click File --> Export
  - Select the BibTeX format and give your file a name
  - choose an export location (the folder with your new .Rmd file)

  3. Go back to your .Rmd file and change the YAML header: write ‘bibliography: library_projecticum.bib’ (or whatever is the name of your .bib file), so your header will look like something this:

```
title: "Untitled" 
output: html_document 
bibliography: library_projecticum.bib
```

  4. Go back to Mendeley, and check the citation key of the random paper we downloaded. You can find it in the detail tab at the right of the screen, scroll down a bit. In my example below, it is `VanLissa2021`

```{r, echo=FALSE, message=FALSE, fig.width=8 }
knitr::include_graphics(here::here(
    "images",
    "mendeley_detailwindow.png"
  )
)
```

Scroll down a bit:


```{r, echo=FALSE, message=FALSE, fig.width=4 }
knitr::include_graphics(here::here(
    "images",
    "mendeley_citationkey.png"
  )
)
```


You can now cite papers it in your .Rmd using citation keys:
  
  ```
  They said something really smart [@VanLissa2021].
  ```
or two people:

  ```
  Some people agree with her [@dehaan08; @vangestel18].
  ```

or include additional information:

  ```
  As do colleagues[see @teunis99, pp. 33-35; also @svanoef04, ch. 1].
  
  or inline:
  
  @smith04 [p. 33] says blah.
  
  ```

Use a minus sign if you don't want to include the names:

  ```
Van der Ploeg says so too [-@vdploeg04].
  ```

</br>
[CC BY-NC-SA 4.0](https://creativecommons.org/licenses/by-nc-sa/4.0/)
