# Working with relational data & databases {#relationaldb}

```{r setup, include=FALSE}
les <- 9
knitr::opts_chunk$set(echo = TRUE)
```

## Resources and references
Part of the materials in this chapter were reproduced in a an adapted form from [@r4ds]. DBeaver tutorials from youtube (see \@ref(resources)). Tutorials on using the package `{RPostgreSQL}` can also be found in \@ref(resources)


## Intro
Relational data is 


## Joining data
 - Joining operations are frequenly used to increase or reduce the amount of data you have
 - To join tables you need a shared variable called a `key`
 - You can have multiple `keys` to join on 
 - There are anumber of `join` verbs
 - We start with the `join` verbs in dplyr
 - We end with join operations using SQL
 - We sometimes which to reshape the format of the data
 - In previous lessons we learned how to do this with the `{tidyr}`package. The functions that we need to recall are: `pivot_longer()`, `pivot_wider()`  
 - Understanding how joins work is essential for working with data stored in relational databases


## Reshaping your data
In order to get your data in the right shape for joining we sometimes need to reshape it. Usually, having or transforming your data in the `tidy` shape is a good practice if you want to perform subsequent joins.

## Remember `tidy` data

 1. All variables in their own column
 1. All observations in their own row
 1. Every cell contains a single observations
 
Below are two illustrations on how to transfer wide to long and vice versa. These examples can be used to understand the workings of the functions `pivot_longer()` and `pivot_wider()`
```{r}
library(dataAnim)
#devtools::install_github("chrk623/dataAnim")

data(datoy_wide)
gather_anim(key = "Subject", value = "Score", col = c("English", "Maths"), data = datoy_wide)

spread_anim(key = "Subject", value = "Score", data = datoy_long)

```

## Relational data




## Creating a PostgreSQL database in DBeaver
In order 




## Using R to work with PostgreSQL databases
On the server th


##### Exercise `r les` {-}
## <mark> EXERCISE 1 </mark> {-} 
 
**TIPS**

 - Be aware, the flu and dengue data contains metadata that should be stripped from the data on load.
 - Think of a way to create valid country names that fit with the gapminder data.
 - Use `readr::write_csv()` and `readr::write_rds()` to write cleaned tables to disk
 
 a) Load the flu ("./data/flu_data.csv), the dengue ("./data/dengue_data.csv) and the gapminder (`{dslabs}` package) into three separate dataframes in R
 
 b) Check if they are in the right shape. Is the data in the 'tidy' format? If not change the format to 'tidy' 
 
 c) Change the `country` and `date` variables of the three tables so that they coincide in terms of data type, class and values 
 
 d) Store the three tables as seperate (so six in total) .csv and .rds files.



    In R, maak van de wide tables een long/stacked formaat met {dplyr} (pivot longer()). Sla de nieuwe tabellen op als rds (R binary)
    In Dbeaver creeer een nieuwe lege PostgreSQL database "workflowsdb"
    Met behulp van R script ({RPostgreSQL} insert de twee tabellen in de "workflowsdb" database
    Inspecteer de content van de tabellen met SQL (in DBeaver) en in R met behulp van {dplyr} en {dbplyr}
    Laadt de gapminder data en pas de data in R zo aan dat een join met de dengue en flue data in de databse mogelijk is (bijvoorbeeld landnamen hetzelfde maken)
    Sla deze clean gapminder data op in de "workflowsdb" database
    Doe een aantal joins en inspecties op de data met SQL (bijvoorbeeld in DBeaver of met dplyr.
    Genereer een grote tabel joined en exporteer die van de database naar de R sessie
    Maak een aantal visualisatie op basis van de joined data met ggplot2 etc...


## Portfolio assignment `r paste0(les, " ")` {-}
