# Masterclass 1 - Shiny, getting started {#shinygettingstarted}

## Aim of this lesson

This lesson is about building Shiny apps to create interactive visualizations, apps and dashbords. We will build a small Shiny app and Dashboard to get you started with `{shiny}`. Thirdly, we will also have a look at how to increase interactivity even more by creating interactive visualizations.

## Resources

[`{flexdashboards}`](https://rmarkdown.rstudio.com/flexdashboard/)

[`{shiny}`](https://shiny.rstudio.com/)

[shiny webinars](https://shiny.rstudio.com/tutorial/)

[`{shinydashboard}`](https://rstudio.github.io/shinydashboard/get_started.html)

## Shiny
Adapted from: https://shiny.rstudio.com/tutorial/
Shiny is an R package that makes it easy to build interactive web applications (apps) straight from R. This lesson will get you started building Shiny apps, though a set of simple examples. Than you need to create a new shiny app from covid-19 case data that is supplied by the RIVM and reflects the covid-19 cases on the city level. 

## My first shiny
To get started with shiny you will need the `{shiny}` R package. The shiny package has a number of build-in isolated shiny apps as examples.
```{r, eval=FALSE}
# install.packages("shiny")
library(shiny)
runExample("01_hello")
```
This will show a histogram of the famous 'faithful' dataset in a pop-up window. The slider can be used to generate multiple versions of this histogram with different number of bins.
So a shiny app is in principle a reactive piece of R code that is run from the input the user specifies. 

##### <div class="question">
Exercise 4.5

Go over the following steps to complete this exercise  

 1. Inspect the source code of the build-in shiny package example "01_hello" that we ran above. 
 1. Copy the source code to an R script (call this file `app.R`) that you store in an `app_example_01_hello` folder in your portfolio project.
 1. Try running the `app.R` file from your portfolio project (open a new session in RStudio)
 1. Now examine the source code and try adjusting the number of bins that is diplayed in the app by default, try a number of different settings
 1. Which piece of code regulates this setting? In which part of the app (`ui` or `server`) is this managed?

</div>

### Interactivity 
The responsiveness of an R object inside a shiny app is a core feature of most shiny apps. This means that for example visualizations based on data are immediatly changed if the user packs a different setting in the UI. This reposiveness can be addedd to an app by using the `reactive()` function in the `{shiny}` package. By using a reactive dataframe for example the input data that is used for example for a table of a graph is updated as soon as the user picks a different setting in the app. Using `reactive()` can greatly speed up the performance of an app.   

We can see the `reactive()` function in action in the following example
```{r, eval=FALSE}
runExample("03_reactivity")
```
Look at the app source code to understand what the `reactive()` part of the code does in this case.


##### <div class="question">
Exercise 4.5

Go over the following steps to complete this exercise  

 1. Study the source code of the build in shiny app example called "04_mpg"
 1. Look at the overall app structure, what is the input and output? Explain in your own words
 1. Look at the `reactive()` part of the server code, what does it do?

<details><summary>Click for the answer</summary>
```{r, eval=FALSE}
## to access the example needed for this exercise
library(shiny)
runExample("04_mpg")
```

</details>

##### <div class="question">
Assignment 

Complete the following steps to finish this assingment. Upload the resulting `app.R` file in your portfolio github project.

Go over the shiny _Written_ tutorials lesson 1 to 7: https://shiny.rstudio.com/tutorial/. You will need the information in these lessons to complete this assignment.

These are the steps:

 1. Fork and clone [this Github repo](https://github.com/uashogeschoolutrecht/covid_demo) to your Rstudio environment.
 1. Study the source code of `./app/shiny/app.R` file carefully
 1. Study the source code of the `helpers.R` files in the `./app/shiny` folder
 1. Run the app locally, play around with the select input box and see what happens
 1. The app does not display data for the vaccination for some countries and continents, because this data is lacking from the input data. Look around on the internet to see if you can find data that has this information.
 1. How would you incorporate the new data source you found in the previous step? 
 1. If you found a relevant data source: add a piece of code to the app that displays this new data. Update the app.R code 
 1. If you did not find any relevant data to add: rewrite the app so that it displays a notification that "no data on vaccinations is available for the selected country"
 1. As you can see from the code, the current version of the app does not use the `reactive()` function. Rewrite part of the app so that the `reactive()` function is used to generate the input data. 
 1. Once you are done, push your changes to your forked branch of the repo and create a pull request.
 1. Publish your version of the `app.R` (and `helpers.R`) scripts in your portfolio.
 
</div>

