---
params:
  geo_id: NL BE
  month: February March April May June July August September October November December
  year: 2020
---
# Lesson 5 - Part 2 - Advanced RMarkdown {#rmarkdownadvanced}

## Lesson Contents

 - Advanced Markdown syntax (images, links, references, cross-references)
 - Bibtex
 - Parameters (params in yaml header, params in chunks)
 - Using Shiny interactive elements in RMarkdown
 - Using `rmarkdown::render()` to render one or multiple paramterized reports
 - Deployment / setting up Github pages for your portfolio
 - Automating deployment - github actions
 - Bookdown / blogdown
 
 ## Resources

 - https://github.com/r-lib/actions/blob/master/examples/README.Rmd
 - https://rmarkdown.rstudio.com/developer_parameterized_reports.html%23parameter_types%2F
 - https://bookdown.org/yihui/rmarkdown/parameterized-reports.html
 



## Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE
  
  )
```

```{r, include=FALSE}
## Packages
library(utils)
library(tidyverse)
library(tools)
library(glue)
library(readxl) 
library(httr) 
library(zoo)
library(blscrapeR)
```


## Data
```{r, echo=TRUE}
## Read data from public url to csv file
data <- readr::read_csv(
  "https://opendata.ecdc.europa.eu/covid19/casedistribution/csv/data.csv", 
  na = c("", " "))
data
```

## Data cleaning
```{r}

## Filter for `r params$month` and `r params$geo_id`
## create a conversion table to go from month name to month integer
convert_months <- tibble(
  month_name = c(
    "january",
    "february",
    "march",
    "april",
    "may",
    "june",
    "july",
    "august",
    "september",
    "october",
    "november",
    "december"),
  month_number = c(1:12))

  m_names <- str_split(tolower(params$month), pattern = " ") %>% unlist()


  month_id <- convert_months %>%
    dplyr::filter(
      month_name %in% m_names) %>% 
    dplyr::select(month_number)
  
  id <- month_id$month_number

geo_ids <- str_split(string = params$geo_id, pattern = " ") %>% unlist()

data_filter <- data %>%
  dplyr::filter(
    month %in% id,
     geoId %in% geo_ids
)
  
char_col <- map_lgl(
  .x = data_filter,
  .f = is.numeric 
)

#map(
#  .x = data_filter[, !char_col],
#  .f = unique
#)

country_names <- data_filter$countriesAndTerritories %>% unique() %>%
  paste(collapse = ", ")

Caps <- function(x) {
s <- strsplit(x, " ")[[1]]
paste(toupper(substring(s, 1,1)), substring(s, 2),
sep="", collapse=" ")
}

range <- data_filter$month %>% unique() %>% enframe() %>%
  arrange(value)

month_df <- left_join(range, convert_months, by = c("value" = "month_number")) %>%
  dplyr::mutate(month_name = firstupper(month_name))

min_month <- month_df %>%
  dplyr::filter(value == min(month_df$value))

countries_vec <- str_split(
  country_names, 
  pattern = ",", 
  simplify = FALSE
  ) %>%
  unlist() %>%
  trimws() %>%
  blscrapeR::firstupper()  
  

max_month <- month_df %>%
  dplyr::filter(value == max(month_df$value))

cap_cases <- paste(
  "COVID-19 positive cases from", 
  min_month$month_name, "to", max_month$month_name,
  "for", glue_collapse(countries_vec, ", ", last = " and ")
)

cap_deaths <- paste(
  "COVID-19 related deaths from", 
  min_month$month_name, "to", max_month$month_name,
  "for", glue_collapse(countries_vec, ", ", last = " and ")
)


data_filter  
```

## Cases
```{r, fig.cap=cap_cases}
set.seed(123)
data_filter <- data_filter %>% 
  mutate(date_time = lubridate::dmy(dateRep))  %>%  # Moving average
  mutate(cases_ma = zoo::rollmean(cases, k = 7, fill = NA))


#names(data_filter)
plot_cases <- data_filter %>%
  na.omit() %>%
  ggplot(aes(x = date_time, y = cases_ma)) +
  geom_point(aes(colour = geoId)) +
  geom_line(aes(group = geoId, colour = geoId)) +
  xlab("Date") +
  ylab("COVID-19 positive tested cases") +
  theme_bw()

plot_cases



```

## Deaths
```{r, fig.cap=cap_deaths}
set.seed(123)

data_filter <- data_filter %>% 
  mutate(date_time = lubridate::dmy(dateRep)) %>%  # Moving average
  mutate(deaths_ma = zoo::rollmean(deaths, k = 7, fill = NA))

#names(data_filter)
plot_deaths <- data_filter %>%
  ggplot(aes(x = date_time, y = deaths_ma)) +
  geom_point(aes(colour = geoId)) +
  geom_line(aes(group = geoId, colour = geoId))  +
  xlab("Date") +
  ylab("COVID-19 related deaths") +
  theme_bw()

plot_deaths

```

## Interactive plot
```{r}

```




