# Portfolio & exam {#portfolioexam}
```{r setuples, include=FALSE}
les <- "setup portfolio"
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## <mark>Plagiarism and copying work of others</mark>
**It is not allowed to hand in a pre-existing analysis as your own work. It is allowed to let yourself be inspired by existing work and copy small pieces of code from pre-existing work, as long as you provide credit to the work you are using or changing. Pay respect to the associated licence if present. It is not allowed to let somebody else create your portfolio for you. All suspects or assessments of fraude or plagiarism will be reported to the Examcommittee of the Institute for Life Sciences & Chemistry**

## Details on exam for this course
As explained, the course will be completed with an assessment on your portfolio. 
The portfolio needs to be hosted on a publicly accessible Gihub pages website. During a live assessment, the teacher and you will go over the portfolio together. The idea is that you provide clarifications on any questions that might arise 
from your portfolio. Two teachers will be present during this assessment.

## Assessment duration
The assessment will be planned during a 20 minutes online video call. Enabling 
your camera during this call is obligatory. The call will start on the planned 
time, so be there on time. Showing up too late in the call without further notice will fail you the assessment. In case you fail the assessment, a new assessment 
will be planned as a do-over. If you fail that do-over again, we will register a 'fail' for the course.

## Portfolio ready before assessment
Make sure that you have shared the public weblink to your portfolio **at least 3 days before the planned assessment**. WE need to have some time in advance to prepare. If, for some reason, you cannot make this deadline, let us know.

## Getting help
We are here to help you pass this course! If you need assistence on getting your portfolio in order, please let us know. We do not have all the answers though, sometimes we will merely be able to point you in a solutions direction. You can 
get help also on fora, either on public fora such as Stack Overflow. Or on the course forum on Canvas. I can highly recommend posing a question on Stack Overflow yourself. If you stick to the forum rules, you will find that there are a lot of friendly fellow-programmers willing to aid you in your programming quest. 

## Setting up a bookdown project and publishing it online as course portfolio {#bookdownportfolio}

The part below will show you how to:

 1. Setup github pages
 1. Create a bookdown project
 1. Render a bookdown project to a local website
 1. Add the rendered site to your github pages
 
## Github pages
Each account on github can be associated to one or multiple websites that are hosted on Github.com. To setup a website on Github (also called `github-pages` or `gh-pages`), we need to create a repository with a special name. For this course we will require you to upload results from the exercises and assignments results to you personal github pages. **This is a hard requirement for passing this course. If you have no or an empty portfolio, you cannot pass this course**

<div class="question">
##### Exercise `r les`

 1. Execute the steps explained [here](https://guides.github.com/features/pages/) to setup a github-pages repository called `workflows-portfolio`
 1. Send the url to your public portfolio address to the teachers by personal chat. 

</div>

## Bookdown
The `{bookdown}` R package is basically a collection of RMarkdown files, collated into a website with nice formatting and the possibility to cross reference and add citations and visualizations. The author of the bookdown is also the main author of `{knitr}`.    

## A simple bookdown example
To start with bookdown, there is a [good and free resource available](https://bookdown.org/yihui/bookdown/get-started.html). 

## Transform your current portfolio project to a bookdown site
By now you should have collected several Rmd files for completing the exercises and assignments in the course thusfar. If everything went well, you have these files collected in your portfolio RStudio project, linked to a Github repo.
To start building your bookdown project, use the minimal example you have seen above. Start by rendering your project using `bookdown::render()`. To the very least you should have an index.html file that contains the yaml definitions for your bookdown project. You can arrange the chapters by using an index in every Rmd file. Use indexes like `001_<file_name>.Rmd`. When you want to add a chapter in between two chapters, this makes it easier. to achieve this.

## Render the book locally
When you have build your `index.Rmd` file and have collected all files in the root of your project, be sure all depending data, scripts and other files such as images etc. are also present in the project. Use the guerrilla analytics logic to setup the project structure. Set the option `new_session: yes` in your `_bookdown.yml` or `index.Rmd` file. This prevent problems down the line. Render the book until all errors are solved. You can debud each file and error by looking at the line numbers where the rendering is causing problems.

## Hosting the bookdown project in your github pages

**IMPORTANT NOTE ON CONFIDENTIALITY**

<mark>**When hosting you portfolio on a public web address, please make sure that there is no confidential information visable in your project. Data that falls under a non-disclosure agreement for example from your projecticum may not be shared. Also, be sure not to publish any personal secrets such as passwords and tokens, inside a github repo or webpage.**</mark> 

Once you have sucefully rendered your bookdown to a full website, hosting can be achived by copying all html files to your Github-pages. The easiest way is to clone your github pages repo and add, commit and push all html files to your github pages repo. The website will be automatically updated after some time. You can also setup a theme for your github pages. See [here](https://docs.github.com/en/github/working-with-github-pages/adding-a-theme-to-your-github-pages-site-with-the-theme-chooser) for more info.

For a complete manual on working with github pages, [see the gh-pages docs](https://docs.github.com/en/github/working-with-github-pages/getting-started-with-github-pages) 

If you want to go for a full continuous integration workflow (which is highly recommended), you can follow these steps in [this blog](https://medium.com/@delucmat/how-to-publish-bookdown-projects-with-github-actions-on-github-pages-6e6aecc7331e)


