# Lesson 7 - RMarkdown - Output formats {#rmarkdownoutput}
```{r setuples, include=FALSE}
les <- 7
```

## Lesson aim
After this lesson you will be able to

 1. have an overview of different output formates for RMarkdown and render those formates.
 1. Setup github pages
 1. Create a bookdown project
 1. Render a bookdown project to a local website
 1. Add the rendered site to your github pages

## Output possibilities

With Rmarkdown you can generate many different things. This reader was made in Rmarkdown. The DAUR1 capstone pdf? RMarkdown. The latest report Alyanne had to send to an ethical committee? Again RMarkdown. Every time we send you all very similar but slightly different emails with stuff like your server password? We generate those using Rmarkdown.

These websites are examples too: [click](https://rmarkdown.rstudio.com/flexdashboard/), [click](https://www.r-graph-gallery.com/index.html).

Have a look around on the following website to get a feeling for what is possible with Rmarkdown.

- [https://r-markdown-gallery.com/index.html](https://r-markdown-gallery.com/index.html)
- [https://bookdown.org/yihui/rmarkdown/basics-examples.html](https://bookdown.org/yihui/rmarkdown/basics-examples.html)

The options include:

  - documents
    - pdf
    - Word
    - html (also: blogs)
    - handouts
    - books
  - interactive documents
    - html widgets
    - dashboards
    - shiny applications
  - presentations
    - slides (Beamer, ioslides, Slidy etc)
    - Powerpoint presentation

### Setting output format

You can specifically choose which output format to render to with the RStudio "Knit" button in the toolbar, or with:

```
rmarkdown::render("your_file_name.Rmd", output_format = "word_document")
```

But usually, you use the YAML header to control which output format(s) knitting will produce and even customize the options for each format separately.
To customize each output format, you can alter the YAML header output argument:

```
---
title: "My document"
output:
  html_document:
---
```

You can change **html_document** to for instance **pdf_document** or **word_document**.

Or customize HTML and PDF output both, this is useful if you want to render multiple types of output.:

```
---
title: "My document"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    fig_caption: false
---
```

<div class="question">

##### Exercise `r les` {-}

[exercise from r4ds here](https://r4ds.had.co.nz/r-markdown.html#exercises-71)

Create one new R Markdown document for each of the three built-in formats: HTML, PDF and Word. Knit each of the three example documents. How does the output differ? How does the input differ? (You may need to install LaTeX in order to build the PDF output — RStudio will prompt you if this is necessary.)
</div>

### Output options

The YAML header can contain a lot more settings for your output. For instance, you may want to add a table of contents (TOC) to you html page, in this example a floating one:

```
---
title: "My document"
output:
  html_document:
    toc: true
    toc_float: true
---
```

Other settings can just be added below html_document:

#### Tabbed sections {-}

You can organize content using tabs by applying the .tabset class attribute to headers within a document. This will cause all sub-headers of the header with the .tabset attribute to appear within tabs rather than as standalone sections. For example:

```
## Quarterly Results {.tabset}

### By Product

(tab content)

### By Region

(tab content)
```

#### Themes {-}

One nice setting is applying themes. You learned how to specify you own formatting with css last lesson. But RMarkdown has some [preset themes](https://www.datadreaming.org/post/r-markdown-theme-gallery/). And of course, you can download packages with more themes.

---
title: "Stuff"
output:
  html_document:
    theme: united
---



<div class="question">

##### Exercise `r les` {-}

  A) What is a floating TOC?
  B) Download Lorem.Rmd [here](XXX) and add a TOC
  C) Gove the TOC a set depth (toc_depth) and see what happens if you change it. (for instance `toc_depth = 2` can be set just as you set `toc_float: true` earlier.)
  D) Also try `code_folding: hide` and `number_sections: FALSE`.
  E) Google how to put two authors on one .Rmd, with affiliations. Put your and your labpartners name as authors.
  F) Apply a theme you like, but not "united".
  G) try tabbed sections

</div>

 
## Interactivity {#rmarkdowninteractive}



## PDF 

XXX PDF themes.
quirks


## Bookdown {#bookdownportfolio}

XXX intro bookdown

You have generated a couple of .Rmd files for your portfolio by now. The next task will be to bundle them into one portfolio website, hosted by github.

We will be setting up a bookdown project and publishing it online as course portfolio 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


 
## Github pages
Each account on github can be associated to one or multiple websites that are hosted on Github.com. To setup a website on Github (also called `github-pages` or `gh-pages`), we need to create a repository with a special name. For this course we will require you to upload results from the exercises and assignments results to you personal github pages. **This is a hard requirement for passing this course. If you have no or an empty portfolio, you cannot pass this course**

<div class="question">
##### Exercise `r les`

 1. Execute the steps explained [here](https://guides.github.com/features/pages/) to setup a github-pages repository called `workflows-portfolio`
 1. Send the url to your public portfolio address to the teachers by personal chat. 

</div>

## Bookdown
The `{bookdown}` R package is basically a collection of RMarkdown files, collated into a website with nice formatting and the possibility to cross reference and add citations and visualizations. The author of the bookdown is also the main author of `{knitr}`.    

## A simple bookdown example
To start with bookdown, there is a [good and free resource available](https://bookdown.org/yihui/bookdown/get-started.html). 

## Transform your current portfolio project to a bookdown site
By now you should have collected several Rmd files for completing the exercises and assignments in the course thusfar. If everything went well, you have these files collected in your portfolio RStudio project, linked to a Github repo.
To start building your bookdown project, use the minimal example you have seen above. Start by rendering your project using `bookdown::render()`. To the very least you should have an index.html file that contains the yaml definitions for your bookdown project. You can arrange the chapters by using an index in every Rmd file. Use indexes like `001_<file_name>.Rmd`. When you want to add a chapter in between two chapters, this makes it easier. to achieve this.

## Render the book locally
When you have build your `index.Rmd` file and have collected all files in the root of your project, be sure all depending data, scripts and other files such as images etc. are also present in the project. Use the guerrilla analytics logic to setup the project structure. Set the option `new_session: yes` in your `_bookdown.yml` or `index.Rmd` file. This prevent problems down the line. Render the book until all errors are solved. You can debud each file and error by looking at the line numbers where the rendering is causing problems.


## Hosting the bookdown project in your github pages

**IMPORTANT NOTE ON CONFIDENTIALITY**

<mark>**When hosting you portfolio on a public web address, please make sure that there is no confidential information visable in your project. Data that falls under a non-disclosure agreement for example from your projecticum may not be shared. Also, be sure not to publish any personal secrets such as passwords and tokens, inside a github repo or webpage.**</mark> 

Once you have sucefully rendered your bookdown to a full website, hosting can be achived by copying all html files to your Github-pages. The easiest way is to clone your github pages repo and add, commit and push all html files to your github pages repo. The website will be automatically updated after some time. You can also setup a theme for your github pages. See [here](https://docs.github.com/en/github/working-with-github-pages/adding-a-theme-to-your-github-pages-site-with-the-theme-chooser) for more info.

For a complete manual on working with github pages, [see the gh-pages docs](https://docs.github.com/en/github/working-with-github-pages/getting-started-with-github-pages) 

If you want to go for a full continuous integration workflow (which is highly recommended), you can follow these steps in [this blog](https://medium.com/@delucmat/how-to-publish-bookdown-projects-with-github-actions-on-github-pages-6e6aecc7331e)


